Smart Spreading Factor Assignment for LoRaWANs

### 2019-01-04 paper.tex
grammer

\begin{abstract}
Low power wide area network (LPWAN) technologies offer affordable connectivity to massive number of low-power devices distributed over very large geographical areas using license-free frequency bands. Focus on this work is one of the most promising LPWAN technologies: LoRa. LoRa offers long range communication and strong resilience to interference by proprietary modulation technique based on Chirp Spread Spectrum. LoRa trades data rate for sensitivity and communication range by spreading symbols within a fixed channel bandwidth. Spreading factor (SF) assignment of nodes has strong effect on collisions thus network performance. In this work, we implemented a simulation environment to study different SF assignment schemes and we proposed a machine learning technique to optimize SF assignment. Finally, we illustrate simulation results for proposed machine learning SF assignment technique.
\end{abstract}

@@

\section{Introduction}
\par In the last few years, number of Internet of Things applications increase exponentially \cite{7721743}. Recent developments on LPWAN technologies has great impact on IoT application number growth. LPWAN technologies offer solutions to some of the oldest wireless communication challenges. Traditional wireless communication methods such as cellular networks (e.g., 2G, 3G, LTE) and short-range communication methods (e.g., Bluetooth, WiFi, Zigbee) cannot provide low power and long range at the same time. Cellular networks can provide long range and high data rate but they are complex and consume too much power. Besides, most of the IoT applications don't require high data rate. Short-range communication methods can provide relatively low power consumption but their range is limited to a few hundred meters at best \cite{7815384}. LPWAN technologies fill the technology gap between short range and cellular communication by providing low power and long range communication. LPWAN technologies basically sacrifice data rate to provide low power consumption.

\par There are several emerging LPWAN technologies. LoRa, Sigfox, NB-IoT and LTE-M are commonly used, well known LPWAN technologies. LoRa and Sigfox use license free ISM frequency bands while NB-IoT and LTE-M use licensed frequency bands which brings extra cost \cite{7815384}. Both LoRa and Sigfox known for ultra low power consumption and resilient to interference while NB-IoT and LTE-M are promoted for higher data rate. LoRa has open standard MAC protocol called LoRaWAN. LoRaWAN and Sigfox MAC protocols are based on pure ALOHA medium access.  LoRaWAN networks can be deployed as private network. However, Sigfox and NB-IoT supports only operator contracted deployments \cite{7815384}. Number of messages that Sigfox end device can send in a day is limited to 140 packets for uplink and just 4 packets for downlink. Sigfox packet payload is limited to 12 bytes for uplink and 8 bytes for downlink. However LoRa supports up to 243 bytes payload and NB-IoT supports up to 1600 bytes payload. Sigfox maximum data rate is also very low 100 bps, on the other hand maximum data rate for LoRa and NB-IoT are 50 kbps and 200 kbps respectively \cite{7815384}.

\par LoRa can adjust data rate by spreading symbols within a fixed channel bandwidth. This enables tradeoff between receive sensitivity and transmission air time \cite{7803607}. LoRa supports 6 different SF option. Simultaneous same SF transmissions are prone to collision however different SF transmissions in same channel are orthogonal to each other, thus SF assignment is crucial for overall network performance. In this work, we introduce a custom made LoRa discrete event simulator to study LoRa SF assignment issue. We also proposed an SF assignment scheme to increase network performance with the help of machine learning methods and we verified our proposal using our simulator.

>>>>

\begin{abstract}
Low power wide area network (LPWAN) technologies offer affordable connectivity to massive number of low-power devices distributed over very large geographical areas using license-free frequency bands. Focus of this work is one of the most promising LPWAN technologies: LoRa. LoRa offers long range communication and strong resilience to interference by proprietary modulation technique based on Chirp Spread Spectrum. LoRa trades data rate for sensitivity and communication range by spreading symbols within a fixed channel bandwidth. Collisions in LoRaWAN networks are strongly related with spreading factor (SF) assignment of nodes which indeed effects network performance. In this work, a simulation environment to study different SF assignment schemes is implemented. Furthermore, novel SF assignment strategy is proposed which utilizes a machine learning technique for optimization. Finally, promising simulation results for proposed machine learning SF assignment technique is presented.
\end{abstract}

@@

\section{Introduction}
\par In the last few years, number of Internet of Things (IoT) applications increased exponentially \cite{7721743}. Recent developments on LPWAN technologies has great impact on growth of number of IoT applications. LPWAN technologies addresses some of the well-known wireless communication challenges. Traditional wireless communication methods such as cellular networks (e.g., 2G, 3G, LTE) and short-range communication (e.g., Bluetooth, WiFi, Zigbee) cannot provide low power and long range at the same time. Cellular networks can provide long range and high data rate but they are complex and consume too much power. Besides, most of the IoT applications do not require high data rate. Short-range communication methods can provide relatively low power consumption but their range is limited to a few hundred meters at best \cite{7815384}. LPWAN technologies fill the technology gap between short range and cellular communication by providing low power and long range communication. LPWAN technologies basically sacrifice data rate to provide low power consumption.

\par There are several emerging LPWAN technologies. LoRa, Sigfox, NB-IoT and LTE-M are commonly used, well-known LPWAN technologies. LoRa and Sigfox use license free ISM frequency bands while NB-IoT and LTE-M use licensed frequency bands which brings extra cost \cite{7815384}. Both LoRa and Sigfox are known for ultra low power consumption and resilient to interference while NB-IoT and LTE-M are promoted for higher data rate. LoRa has open standard MAC protocol called LoRaWAN. LoRaWAN and Sigfox MAC protocols are based on pure ALOHA medium access. LoRaWAN networks can be deployed as a private network. However, Sigfox and NB-IoT are only available with operator contracted deployments \cite{7815384}. Number of messages that Sigfox end device can send in a day is limited to 140 packets for uplink and just 4 packets for downlink. Also, Sigfox packet payload is limited to 12 bytes for uplink and 8 bytes for downlink. However LoRa supports up to 243 bytes payload and NB-IoT supports up to 1600 bytes payload. Sigfox maximum data rate is 100 bps, on the other hand maximum data rate for LoRa and NB-IoT are 50 kbps and 200 kbps respectively \cite{7815384}.

\par LoRa can adjust data rate by spreading symbols within a fixed channel bandwidth. This enables tradeoff between receive sensitivity and transmission time on air \cite{7803607}. LoRa supports 6 different SF option. Simultaneous same SF transmissions are prone to collision however different SF transmissions in same channel are orthogonal to each other up to some extent, thus SF assignment is crucial for overall network performance. In this work, a custom made LoRa discrete event simulator is introduced to study different LoRa SF assignment strategies. We also proposed an SF assignment scheme to increase network performance with the help of machine learning methods and we verified our proposal using the simulator.
